---
title: "Soccer Preformance Analysis"
author: 
- "Kelly Rivera"
- "Contribitors: Rachel Tuffy, Manuela Montoya, Dr. Sean Mondesire"
output: html_notebook
---

# Import Libraries
```{r}
#install.packages("tidyverse")
#install.packages("corrplot")
#install.packages("Metrics")
```


# Load Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(corrplot)
library(Metrics)
```

# Import Data
```{r}
FieldPlayers <- read_csv("FieldPlayers.csv") %>%
  na.omit(Result)
FieldPlayers
```

# Data Preparation

Defenders
```{r}
defenseData <- FieldPlayers[FieldPlayers$`Position Category` == "Defenders", ] 

defenseData$`Total Touches` <- defenseData$`Total Touches (#)`

defenseData
```
Midfield
```{r}
midfieldData <- FieldPlayers[FieldPlayers$`Position Category` == "Midfielders", ] 

midfieldData
```

Forwards
```{r}
forwardsData <- FieldPlayers[FieldPlayers$`Position Category` == "Forwards", ] 

forwardsData$`Releases per min` <- forwardsData$`Releases per min (#/min)`
forwardsData$`RV Zone 1` <- forwardsData$`RV Zone 1 [0-10( m/s)]`
forwardsData$`Avg Time on Ball` <- forwardsData$`Avg Time on Ball - Short (sec)`
forwardsData$`Short Possessions` <- forwardsData$`Short Possessions (#)`

forwardsData
```


# Data Analytics 

Defenders Correlation
```{r}
selectDefenseData <- select(defenseData, Result, 'Total Touches')
corValueD <- cor(selectDefenseData, use = "complete.obs")
corrplot(corValueD, method = "color", addCoef.col = "black", type = "lower")
```

Midfield Correlation
```{r}
selectMidfieldData <- select(midfieldData, Result, 'Top Speed (m/s)', 'Work Rate (m/min)', 'Receives Right (#)')

corValMid <- cor(selectMidfieldData, use = "complete.obs")

corrplot(corValMid, method = "color", addCoef.col = "black", type = "upper", title = "Midfield Correlation Matrix", mar = c(0,0,1,0))
```

Forwards Correlation
```{r}
selectForwardsData <- select(forwardsData, Result, SH, `SOG`, `SOG/Sh`, 'Releases per min',
                             `RV Zone 1`, `One-Touch (#)`, `Short Possessions`, 
                             `Avg Time on Ball`, `Top Speed (m/s)`)

corValueFD <- cor(selectForwardsData, use = "complete.obs")
corrplot(corValueFD, method = "color", addCoef.col = "black", type = "lower", title = "Forwards Correlation Matrix", mar = c(0,0,1,0))
```


# Data Visualiztions 

## Defense graphs
```{r}
ggplot(data = defenseData, mapping = aes(y = `Result`, x = `Total Touches (#)`)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Defenders: Total Number of Touches vs. Match Results",
       subtitle = "Data: 2020 Season",
       x = "Total Number of Touches",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

## Midfield graphs

Result vs Work rate
```{r}
ggplot(data = midfieldData, mapping = aes(y = `Result`, x = `Work Rate (m/min)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Midfielders: Work Rate (m/min) vs. Match Results",
       subtitle = "Data: 2020 Season",
       x = "Work Rate (m/min)",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Receives Right 
```{r}
ggplot(data = midfieldData, mapping = aes(y = `Result`, x = `Receives Right (#)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Midfielders: Number of Receives Right vs. Match Results",
       subtitle = "Data: 2020 Season",
       x = "Number of Receives Right",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Top Speed (m/s) 
```{r}
ggplot(data = midfieldData, mapping = aes(y = `Result`, x = `Top Speed (m/s)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Midfielders: Top Speed (m/s) vs. Match Results",
       subtitle = "Data: 2020 Season",
       x = "Top Speed (m/s)",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

## Forward Graphs
Result vs SH 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `SH`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: SH vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs SOG 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `SOG`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: SOG vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs SOG/Sh 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `SOG/Sh`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: SOG/Sh vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Releases (#) 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `Releases (#)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: Number of Releases vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Releases per min (#/min) 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `Releases per min (#/min)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: Number of Releases per min (#/min) vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs RV Zone 1 [0-10( m/s)] 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `RV Zone 1 [0-10( m/s)]`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: RV Zone 1 [0-10( m/s)] vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs One-Touch (#) 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `One-Touch (#)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: Number of One-Touches vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Short Possessions (#) 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `Short Possessions (#)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: Number of Short Possessions vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```

Result vs Avg Time on Ball - Short (sec) 
```{r}
ggplot(data = forwardsData, mapping = aes(y = `Result`, x = `Avg Time on Ball - Short (sec)`)) +
geom_point() +
geom_smooth(method = "lm") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Forwards: Avgerage Seconds on Ball vs. Match Results",
       subtitle = "Data: 2020 Season",
       y = "Game Result",
       caption = "*3 is equivelent to a Win")
```


# Prediction Models & RMSE
```{r}

# Linear Regression Defense
lmModelDefense <- lm(Result ~ `Total Touches (#)`, data = defenseData)
predictedLM_Defense <- predict(lmModelDefense, data = defenseData)


# Linear Regression Midfield
lmModelMidfield <- lm(Result ~ `Top Speed (m/s)` + `Work Rate (m/min)` + `Receives Right (#)`, data = midfieldData)
predictedLM_Midfield <- predict(lmModelMidfield, data = midfieldData)


# Linear Regression Forwards
lmModelForwards <- lm(Result ~ `SH` + `SOG` + `SOG/Sh` + `Releases (#)` + `Releases per min (#/min)` + `RV Zone 1 [0-10( m/s)]` +
                     `One-Touch (#)` + `Short Possessions (#)` + `Avg Time on Ball - Short (sec)`, data = forwardsData)
predictedLM_Forwards <- predict(lmModelForwards, data = forwardsData)



# Calculate error: Root mean squared error (RMSE)
rmseLM_Defense <- rmse(defenseData$Result, predictedLM_Defense)
rmseLM_Midfield <- rmse(midfieldData$Result, predictedLM_Midfield)
rmseLM_Forwards <- rmse(forwardsData$Result, predictedLM_Forwards)


# Print the RMSE of each position model
cat("\nRMSE:\n")
cat("Defense LM:", rmseLM_Defense, "\n")
cat("Midfield LM:", rmseLM_Midfield, "\n")
cat("Forwards LM:", rmseLM_Forwards, "\n")

```

# Error Reporting
```{r}
errorData <- data.frame('Models' = c("Defender", "Midfielder", "Forward"), 
                 'RMSE' = c(1.293275, 0.8685143, 0.01))
print(errorData)
```

## Error Visualization

```{r}
ggplot(errorData, aes(x = Models, y = RMSE, fill = Models)) + 
    geom_col(show.legend = FALSE) +
  labs(title = "Linear Regression Model Acceracy",
       #caption = "Each model predicted the Match Result given position based variables",
       x = "Model by Position",
       y = "Root Mean Square Error")

```